### 2021-12-05 Stat 400 proj.Rmd
Fixed simulation error, turns out it didnt make sense to see which sus ppl get infected when theres no sus ppl

---
title: "Stat 400 Project"
authors: "Nathan Hemenway, Mark Hinds, Ian Hall"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#testing github
library(tidyverse)

#create starting positions and state function
starting_vals <- function(num_subjects, I_0, x_dim, y_dim){

  #Infect I_0 many starting subjects
  subjects <- rep('S', num_subjects)
  infect <- sample(x = 1:num_subjects, size=I_0, replace=FALSE)
  subjects[infect] <- 'I'

  #Storage for x, y starting positions for the subjects
  start_x <- runif(n=num_subjects, min=0, max=x_dim)
  start_y <- runif(n=num_subjects, min=0, max=y_dim)

  inc_time <- rep(-1,num_subjects)
  rec_days <- rep(-1,num_subjects)

  #Put into a data frame
  df <- as.data.frame(cbind(as.numeric(start_x), as.numeric(start_y), subjects, rec_days, inc_time))
  colnames(df)[1] <- 'x'
  colnames(df)[2] <- 'y'
  colnames(df)[3] <- 'subjects'
  colnames(df)[4] <- 'incubation'
  colnames(df)[5] <- 'recovery'
  return(df)
}

#Random step length function
get_step_length <- function(df, x_dim, y_dim){
  step_lengths <- rnorm(n=nrow(df), mean=min(x_dim, y_dim)/4, sd=min(x_dim, y_dim)/12)
  return(step_lengths)
}

#Random step direction function
get_step_direction <- function(length_vec){
  direction <- runif(n=length(length_vec), min=0, max=2*pi)
}

#Get delta_x, delta_y
get_delts <- function(length_vec, direc_vec){
  delta_x <- length_vec * cos(direc_vec)
  delta_y <- length_vec * sin(direc_vec)
  return(cbind(delta_x, delta_y))
}

#Function for getting updated coordinates
update_coords <- function(og_df, delta_df){
  new_mx <- matrix(data=NA, nrow=nrow(og_df), ncol=5)
  new_df <- data.frame(new_mx)
  new_df[,1] <- as.numeric(og_df[,'x']) + delta_df[,'delta_x']
  new_df[,2] <- as.numeric(og_df[,'y']) + delta_df[,'delta_y']
  new_df[,3] <- og_df[,3]
  new_df[,4] <- og_df[,4]
  new_df[,5] <- og_df[,5]
  colnames(new_df) <- c('x', 'y', 'subjects', 'recovery', 'incubation')
  new_df
}

coord_checker <- function(og_df, new_df, x_dim, y_dim){
  baddies <- NULL
  for(i in 1:nrow(new_df)){
    if(as.numeric(new_df[i, 'x']) > x_dim | as.numeric(new_df[i, 'x']) < 0 | as.numeric(new_df[i, 'y']) > y_dim | as.numeric(new_df[i, 'y']) < 0){
      baddies <- append(baddies, i)
    }
  }

  if(is.null(baddies)){
    return(new_df)
  }else{
  new_step <- get_step_length(new_df[i,], x_dim, y_dim)
  new_direction <- get_step_direction(new_step)
  new_change <- get_delts(new_step, new_direction)
  new_df <- update_coords(og_df[baddies,], new_change)
  full_df <- rbind(new_df, og_df[-baddies,])
  return(coord_checker(og_df, full_df, x_dim, y_dim))
  }
}

#Function for deciding if infections occurred
#BUG -- If the mean_crit_dist is small, all infections go away!!
infection_status <- function(state, mean_crit_dist){

  current_infected <- state[state[,3]=='I',]
  current_suceptable <- state[state[,3]=='S',]
  current_removed <- state[state[,3]=='R',]

  #store new infection status
  newly_infected <- NULL

  #Check to see which susceptible subjects get infected
  #For loop
  if(nrow(current_suceptable) != 0){
  for(i in 1:nrow(current_infected)){
    for(j in 1:nrow(current_suceptable)){
      x_i <- as.numeric(current_infected[i, 'x'])
      y_i <- as.numeric(current_infected[i, 'y'])
      x_s <- as.numeric(current_suceptable[j, 'x'])
      y_s <- as.numeric(current_suceptable[j, 'y'])

      crucial_distance <- rexp(1, 1/mean_crit_dist)
      d <- sqrt((x_i - x_s)^2 + (y_i - y_s)^2)
      # print(c(d <= crucial_distance, x_i, x_s, j, i, nrow(current_infected)))
      if(d <= crucial_distance){
        newly_infected <- c(newly_infected, j)
      }

    }
  }

  current_suceptable[newly_infected, 3] <- 'I'
  #Give incubation time
  current_suceptable[newly_infected, 4] <- rpois(n=length(newly_infected), lambda = 5)
  #Give recovery time
  current_suceptable[newly_infected, 5] <- rpois(n=length(newly_infected), lambda = 14)

  new_state <- rbind(current_infected, current_suceptable, current_removed)
  return(new_state)
  }else{return(state)}
}

#simulation function
simulator <- function(days, num_subjects, x_dim, y_dim, I_0){
  initial_data <- starting_vals(num_subjects = num_subjects, x_dim=x_dim, y_dim=y_dim, I_0=I_0)
  end_day <- list()

  for(i in 1:days){

    if(i == 1){
      step_size <- get_step_length(initial_data, x_dim, y_dim)
      direction <- get_step_direction(step_size)
      change <- get_delts(step_size, direction)

      updated_positions <- update_coords(initial_data, change)
      checked_positions <- coord_checker(og_df=initial_data, new_df=updated_positions, x_dim=x_dim, y_dim=y_dim)
      new_status <- infection_status(checked_positions, 2)

      not_recovered <- new_status[as.numeric(new_status[, 5]) > 0, ]
      new_time <- as.numeric(not_recovered[, 5]) - 1
      not_recovered[, 5] <- new_time
      infected_0 <- new_status[as.numeric(new_status[, 5]) == 0, ]
      infected_0[, 3] <- rep('R', nrow(infected_0))
      not_infected = new_status[as.numeric(new_status[, 5]) < 0, ]

      new_status = rbind(not_recovered, infected_0, not_infected)
      end_day[[i]] <- new_status

    }

    else{
      step_size <- get_step_length(initial_data, x_dim, y_dim)
      direction <- get_step_direction(step_size)
      change <- get_delts(step_size, direction)

      updated_positions <- update_coords(end_day[[i-1]], change)
      checked_positions <- coord_checker(og_df=end_day[[i-1]], new_df=updated_positions, x_dim=x_dim, y_dim=y_dim)
      new_status <- infection_status(checked_positions, 5)

      not_recovered <- new_status[as.numeric(new_status[, 5]) > 0, ]
      new_time <- as.numeric(not_recovered[, 5]) - 1
      not_recovered[, 5] <- new_time
      infected_0 <- new_status[as.numeric(new_status[, 5]) == 0, ]
      infected_0[, 3] <- rep('R', nrow(infected_0))
      not_infected = new_status[as.numeric(new_status[, 5]) < 0, ]

      new_status = rbind(not_recovered, infected_0, not_infected)
      end_day[[i]] <- new_status

    }
  }
  return(end_day[[1]])
}

#Testing
df <- starting_vals(10, 2, 1, 1)
len <- get_step_length(df, 1, 1)
dir <- get_step_direction(len)
delts <- get_delts(len, dir)
newdf <- update_coords(df, delts)
coords <- coord_checker(df, newdf, 1, 1)
coords
new_status <- infection_status(coords, 2)
new_status

x = simulator(100,25,1,1,2)

```

Plots
``` {r}
ggplot() + 
  geom_line(aes())
```

```{r}
#reported monte carlo summary data
qd = c(0,25,50,75)
pos = c(99.3,99.2,98.7,92.7)
dncpv = c(69,53,46,33)
dncpt = c(26,32,43,42)
r0 = c(1.8,1.73,1.8,1.63)
mcsum = cbind(qd,pos,dncpv,dncpt,r0)
colnames(mcsum,c("Quarantine degree [%]",	"Percentage of size %",	"DNC peak value [cases]",	"DNC peaking time [days]",	"R0"))
```