### 2021-12-14 Paper.Rmd
Added some plots and more to methodology

---
title: "Paper"
author: "Nathan Hemenway"
date: "12/12/2021"
output: pdf_document
---

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(boot)
#create starting positions and state function
starting_vals <- function(num_subjects, I_0, x_dim, y_dim, prop_vac){

  #Infect I_0 many starting subjects
  subjects <- rep('S', num_subjects)
  infect <- sample(x = 1:num_subjects, size=I_0, replace=FALSE)
  subjects[infect] <- 'I'

  #Determining if a person is vaccinated
  vac_ppl <- sample(x = 1:num_subjects, size = round(prop_vac * num_subjects), replace = F)
  vac_vec <- rep(F , num_subjects)
  vac_vec[vac_ppl] <- T

  #Storage for x, y starting positions for the subjects
  start_x <- runif(n=num_subjects, min=0, max=x_dim)
  start_y <- runif(n=num_subjects, min=0, max=y_dim)

  inc_time <- rep(-1, num_subjects)
  rec_days <- rep(-1, num_subjects)

  #Put into a data frame
  df <- as.data.frame(cbind(as.numeric(start_x), as.numeric(start_y), subjects, rec_days, inc_time, vac_vec))
  colnames(df)[1] <- 'x'
  colnames(df)[2] <- 'y'
  colnames(df)[3] <- 'subjects'
  colnames(df)[4] <- 'incubation'
  colnames(df)[5] <- 'recovery'
  colnames(df)[6] <- 'vaccine'

  #Assign a recovery time to the infected subjects
  df[df[,3] == 'I', 5] <- rpois(I_0, 14)

  #Assign an incubation time to the infected subjects
  df[df[,3] == 'I', 4] <- rpois(I_0, 5)

  return(df)
}

#Random step length function
get_step_length <- function(df, x_dim, y_dim){
  step_lengths <- rnorm(n=nrow(df), mean=min(x_dim, y_dim)/4, sd=min(x_dim, y_dim)/12)
  return(step_lengths)
}

#Random step direction function
get_step_direction <- function(length_vec){
  direction <- runif(n=length(length_vec), min=0, max=2*pi)
}

#Get delta_x, delta_y
get_delts <- function(length_vec, direc_vec){
  delta_x <- length_vec * cos(direc_vec)
  delta_y <- length_vec * sin(direc_vec)
  return(cbind(delta_x, delta_y))
}

#Function for getting updated coordinates
update_coords <- function(og_df, delta_df){
  new_mx <- matrix(data=NA, nrow=nrow(og_df), ncol=5)
  new_df <- data.frame(new_mx)
  new_df[,1] <- as.numeric(og_df[,'x']) + delta_df[,'delta_x']
  new_df[,2] <- as.numeric(og_df[,'y']) + delta_df[,'delta_y']
  new_df[,3] <- og_df[,3]
  new_df[,4] <- og_df[,4]
  new_df[,5] <- og_df[,5]
  new_df[,6] <- og_df[,6]
  colnames(new_df) <- c('x', 'y', 'subjects', 'incubation', 'recovery', 'vaccine')
  new_df
}

coord_checker <- function(og_df, new_df, x_dim, y_dim){
  baddies <- NULL
  for(i in 1:nrow(new_df)){
    if(as.numeric(new_df[i, 'x']) > x_dim | as.numeric(new_df[i, 'x']) < 0 | as.numeric(new_df[i, 'y']) > y_dim | as.numeric(new_df[i, 'y']) < 0){
      baddies <- append(baddies, i)
    }
  }

  if(is.null(baddies)){
    return(new_df)
  }else{
  new_step <- get_step_length(new_df[i,], x_dim, y_dim)
  new_direction <- get_step_direction(new_step)
  new_change <- get_delts(new_step, new_direction)
  new_df <- update_coords(og_df[baddies,], new_change)
  full_df <- rbind(new_df, og_df[-baddies,])
  return(coord_checker(og_df, full_df, x_dim, y_dim))
  }
}

#Function for deciding if infections occurred
infection_status <- function(state, mean_crit_dist){

  current_infected <- state[state[,3]=='I' & state[,4] <= 0,]
  current_suceptable <- state[state[,3]=='S',]
  current_removed <- state[state[,3]=='R',]
  current_non_contagious <- state[state[,3]=='I' & state[,4] > 0,]

  #store new infection status
  newly_infected <- NULL

  #Check to see which susceptible subjects get infected
  #For loop
  if(nrow(current_suceptable) != 0 & nrow(current_infected) != 0){
  for(i in 1:nrow(current_infected)){
    for(j in 1:nrow(current_suceptable)){
      x_i <- as.numeric(current_infected[i, 'x'])
      y_i <- as.numeric(current_infected[i, 'y'])
      x_s <- as.numeric(current_suceptable[j, 'x'])
      y_s <- as.numeric(current_suceptable[j, 'y'])

      crucial_distance <- rexp(1, 1/mean_crit_dist)
      vaccine_status <- current_suceptable[j, 'vaccine']
      if(vaccine_status == T) {
        vac_eff <- as.logical(rbinom(1,1,.8))
      }
      else {
        vac_eff <- F
      }

      d <- sqrt((x_i - x_s)^2 + (y_i - y_s)^2)
      if(d <= crucial_distance & (vac_eff == F)){
        newly_infected <- c(newly_infected, j)
      }
    }
  }

  current_suceptable[newly_infected, 3] <- 'I'
  #Give incubation time
  current_suceptable[newly_infected, 4] <- rpois(n=length(newly_infected), lambda = 5)
  #Give recovery time
  current_suceptable[newly_infected, 5] <- rpois(n=length(newly_infected), lambda = 14)

  new_state <- rbind(current_infected, current_suceptable, current_removed, current_non_contagious)
  return(new_state)
  }else{return(state)}
}

#simulation function
simulator <- function(days, num_subjects, x_dim, y_dim, I_0, inf_dist, prop_vac){
  initial_data <- starting_vals(num_subjects = num_subjects, x_dim=x_dim, y_dim=y_dim, I_0=I_0, prop_vac = prop_vac)
  end_day <- list()
  num_cases <- rep(NA, days)
  num_cases <- c(I_0, num_cases)
  num_new_cases <- rep(NA, days)

  for(i in 1:days){

    if(i == 1){
      step_size <- get_step_length(initial_data, x_dim, y_dim)
      direction <- get_step_direction(step_size)
      change <- get_delts(step_size, direction)

      updated_positions <- update_coords(initial_data, change)
      checked_positions <- coord_checker(og_df=initial_data, new_df=updated_positions, x_dim=x_dim, y_dim=y_dim)
      new_status <- infection_status(checked_positions, inf_dist)

      #recovery time code
      not_recovered <- new_status[as.numeric(new_status[, 5]) > 0, ]
      new_time <- as.numeric(not_recovered[, 5]) - 1
      not_recovered[, 5] <- new_time
      infected_0 <- new_status[as.numeric(new_status[, 5]) == 0, ]
      infected_0[, 3] <- rep('R', nrow(infected_0))
      not_infected = new_status[as.numeric(new_status[, 5]) < 0, ]
      new_status = rbind(not_recovered, infected_0, not_infected)

      #incubation time code
      infected_people <- new_status[new_status[,3] == 'I',]
      new_inc_time <- as.numeric(infected_people[,4]) - 1
      infected_people[,4] <- new_inc_time
      other_people <- new_status[new_status[,3] != 'I',]
      new_status <- rbind(infected_people, other_people)

      end_day[[i]] <- new_status

      #Calculate number of new cases
      num_cases[i+1] <- sum(new_status[,3]=='I')
      num_new_cases[i] <- sum(new_status[,3]=='I') - I_0

    }

    else{
      step_size <- get_step_length(initial_data, x_dim, y_dim)
      direction <- get_step_direction(step_size)
      change <- get_delts(step_size, direction)

      updated_positions <- update_coords(end_day[[i-1]], change)
      checked_positions <- coord_checker(og_df=end_day[[i-1]], new_df=updated_positions, x_dim=x_dim, y_dim=y_dim)
      new_status <- infection_status(checked_positions, inf_dist)

      not_recovered <- new_status[as.numeric(new_status[, 5]) > 0, ]
      new_time <- as.numeric(not_recovered[, 5]) - 1
      not_recovered[, 5] <- new_time
      infected_0 <- new_status[as.numeric(new_status[, 5]) == 0, ]
      infected_0[, 3] <- rep('R', nrow(infected_0))
      not_infected = new_status[as.numeric(new_status[, 5]) < 0, ]
      new_status = rbind(not_recovered, infected_0, not_infected)

      #incubation time code
      infected_people <- new_status[new_status[,3] == 'I',]
      new_inc_time <- as.numeric(infected_people[,4]) - 1
      infected_people[,4] <- new_inc_time
      other_people <- new_status[new_status[,3] != 'I',]
      new_status <- rbind(infected_people, other_people)

      end_day[[i]] <- new_status

      #Calculate number of new cases
      num_cases[i+1] <- sum(new_status[,3]=='I')
      num_new_cases[i] <- sum(new_status[,3]=='I') - num_cases[i-1]

    }
  }
  return(num_cases)
}

#bootstrap
mean_func <- function(daytuh, idx) {
  mean(daytuh[idx])
}

get_bootstrap_cis <- function(days, num_subjects, x_dim, y_dim, I_0, inf_dist, sample_size, ci_type='perc', prop_vac = 0){
  #Create data frame to store samples
  samps <- matrix(data=NA, nrow=sample_size, ncol=days + 1)
  samps <- data.frame(samps)

  #Create data frame to store confidence intervals
  ci_df <- matrix(data=NA, nrow=days + 1, ncol=3)
  ci_df <- data.frame(ci_df)

  for(i in 1:sample_size){
    samps[i, ] <- simulator(days, num_subjects, x_dim, y_dim, I_0, inf_dist, prop_vac)
  }

  for(i in 1:ncol(samps)){
    #Make sure column has more than one unique value
    check <- (length(unique(samps[ ,i])) != 1)
    if(check){
    boot_obj <- boot(samps[ ,i], mean_func, 2000)
    ci_obj <- boot.ci(boot_obj, conf = .95, type = ci_type)
    ci_df[i, 1] <- ci_obj$percent[4]
    ci_df[i, 2] <- ci_obj$t0
    ci_df[i, 3] <- ci_obj$percent[5]
    }else{ci_df[i, ] <- c(samps[1, i], samps[1, i], samps[1, i])}
  }
return(ci_df)
}

```

Background:
The paper "Novel approach for Monte Carlo simulation of the new COVID-19 spread dynamics aim's to investigate the transmission and incubation random processes of COVID-19. The goal of the study is to generate epidemiological data based on the natural mechanism's of transmission of the virus assuming random interactions of a large-finite number of individuals in very short distance ranges. A few important assumptions that were made were in how the transmission of the virus occurs when the physical distance between individuals decreases to where the probability of transmission converges to one, asymptotically. Their model defined the steps of movement in a defined space as a randomly assigned process. They approached this through the SIR model; A model defining Susceptible, Infected, Recovered individuals. All change in the SIR model is based on time, as the function iterates through a set number of days. They started with a 0.5% infection rate and simulation daily infection counts over a range of time. In order to capture many factors in the transmission including the relationship between infection probability between physical distance, effective and ineffective mask wearing, and environmental conditions they distributed these factors Exponential(rate = lambda). Initial recovery and incubation times were distributed Gamma(alpha,beta).

Motivation:
The motivation behind our Monte Carlo Covid-19 simulator is to be able to better understand the spread of Covid-19. Through Monte Carlo simulation, we can observe what affect variables such as starting infected population, total population, physical space, infection distance, as well as incubation and recovery times have on the long term spread of the disease. These variables can be easily adjusted in our code to give different scenarios. The way we measure the effect changes in the aforementioned variables is by looking at the difference between the daily number of cases. This can be well visualized in a plot giving number of infected people per day. At the beginning of the Covid-19 pandemic, there were public health measures put into place that aimed at 'flattening' the curve of daily cases of Covid-19. We are interested in seeing what effect different policies and population characteristics have on the plot of daily Covid-19 infections. If we can understand what affect different input variables have on the outcome of the pandemic, we can possibly make better policy decisions regarding Covid-19.

Methodology:
To start, we created functions that generate a random step length. This was done using the rnorm() function for the normal distribution with a mean and standard deviation of the minimum of the dimensions divided by four and twelve respectively. Once we have a step length we generate a step direction from a uniform(0, 2$\pi$) distribution using runif(). This gives the direction angle in radians. From these two pieces of information, we can then update the coordinate values in the update_coords function by adding the change to the original x and y values. This is done using trigonometry equations for the legs of a triangle given the hypotenuse and angle.

The obvious problem that arises when using this approach is that subjects will inevitably leave the specified space they are supposed to remain in. For example, if we want the subjects to stay in a [0, 1] x [0, 1] square, and we let them move in any direction, the mean step distance will be 0.25, so it would only take a centered subject a couple moves in the same direction before they leave the space. The solution we employed was to check and see whether a certain move would put a subject out of bounds, and if so re-sample a different move, then repeat for all subjects until they are all within bounds. This is what the coord_checker() function does.

After moving the subjects randomly in the given space, the next step in the simulation is to check if any infected subjects have infected others. This is done via the infection_status() function. The first thing we did was subset the subjects into different data frames by infection status. Then we can check how far the susceptible subjects are from the infected subjects. If they come within a crucial distance generated using an exponential distribution, then they are vulnerable to infection. Then if the susceptible person is vaccinated, we generate a binomial(1, 0.8) random variable to determine if the vaccine was effective in preventing infection. The third condition necessary for infection to occur is that the incubation period of the virus has passed for the infected person. The incubation length is assigned upon infection using rpois(1, 5), and decreases by one each day until it reaches zero. If binomial vaccine effect returns false, and the susceptible person is within the generated crucial distance of the infected person, and the infected person is past the incubation period, then the suceptible person becomes infected. Once infected, a subject is assigned a recovery time using rpois(1, 14). For each iteration in the simulator, one is subtracted from the recovery time. Once recovery time reaches zero, the infection status of the subject is changed from infected to recovered.

Results:
Going to look at 1x1 with 1000 people: represents small town population density with 0.5% intial infected. Will look at distances between 0.001 (6 ft) to 0.01 (60 ft?) and see what differences exist. Assuming no vaccinations.
```{r, cache=TRUE}
set.seed(400)
#~6ft apart
y <- simulator(150, 1000, 1, 1, 5, 0.001, 0)
x <- 0:150

ggplot() +
  geom_line(aes(x=x, y=y)) +
  xlab('Day') +
  ylab('Number of Cases') +
  ggtitle('Number of Covid-19 cases over time')

#~26 ft apart
y <- simulator(150, 1000, 1, 1, 5, 0.005, 0)
x <- 0:150

ggplot() +
  geom_line(aes(x=x, y=y)) +
  xlab('Day') +
  ylab('Number of Cases') +
  ggtitle('Number of Covid-19 cases over time')

#~ 53ft apart
y <- simulator(150, 1000, 1, 1, 5, 0.01, 0)
x <- 0:150

ggplot() +
  geom_line(aes(x=x, y=y)) +
  xlab('Day') +
  ylab('Number of Cases') +
  ggtitle('Number of Covid-19 cases over time')

```
Write about difference between different crucial distances here.

```{r}
set.seed(400)
plot_dat1 <- get_bootstrap_cis(days=50, num_subjects=200, x_dim=1, y_dim=1, I_0=1, inf_dist=0.1, sample_size=20, prop_vac = 0)

ggplot() +
  geom_line(aes(x=1:51, y=plot_dat1[ ,1]), linetype = 2, color = 'red') +
  geom_line(aes(x=1:51, y=plot_dat1[ ,2])) +
  geom_line(aes(x=1:51, y=plot_dat1[ ,3]), linetype = 2, color = 'red')

plot_dat2 <- get_bootstrap_cis(days=50, num_subjects=200, x_dim=1, y_dim=1, I_0=1, inf_dist=0.1, sample_size=20, prop_vac = 0.5)

ggplot() +
  geom_line(aes(x=1:51, y=plot_dat2[ ,1]), linetype = 2, color = 'red') +
  geom_line(aes(x=1:51, y=plot_dat2[ ,2])) +
  geom_line(aes(x=1:51, y=plot_dat2[ ,3]), linetype = 2, color = 'red')

plot_dat3 <- get_bootstrap_cis(days=50, num_subjects=200, x_dim=1, y_dim=1, I_0=1, inf_dist=0.1, sample_size=20, prop_vac = 0.8)

ggplot() +
  geom_line(aes(x=1:51, y=plot_dat3[ ,1]), linetype = 2, color = 'red') +
  geom_line(aes(x=1:51, y=plot_dat3[ ,2])) +
  geom_line(aes(x=1:51, y=plot_dat3[ ,3]), linetype = 2, color = 'red')

plot_dat4 <- get_bootstrap_cis(days=50, num_subjects=200, x_dim=1, y_dim=1, I_0=1, inf_dist=0.1, sample_size=20, prop_vac = 0.95)

ggplot() +
  geom_line(aes(x=1:51, y=plot_dat4[ ,1]), linetype = 2, color = 'red') +
  geom_line(aes(x=1:51, y=plot_dat4[ ,2])) +
  geom_line(aes(x=1:51, y=plot_dat4[ ,3]), linetype = 2, color = 'red')

#plot all vaccine curves
ggplot() +
  geom_line(aes(x=1:51, y=plot_dat1[ ,2])) +
  geom_line(aes(x=1:51, y=plot_dat2[ ,2]), color='red') +
  geom_line(aes(x=1:51, y=plot_dat3[ ,2]), color='blue') +
  geom_line(aes(x=1:51, y=plot_dat4[ ,2]), color='green')

```